---
title: "Penguin Flipper Lengths ANOVA"
author: "Geoffrey Cook"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load necessary packages and read in the necessary data

```{r}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
View(penguins)
```

Our question: Is there a difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipelago?

The type of test to compare means accross more than 2 groups for one indipendant variable...
ONE-WAY ANOVA

- Independent observations
- Equal variance (Levene's test)
- Normality (histograms/qqplots)

```{r}

counts <- with(penguins, table(species, sex))
View(counts)

# Let's do some updating!

penguins2 <- penguins %>% 
  mutate(
    common_name = 
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-species) %>% # REMOVES a particular column
  filter(sex == "FEMALE")
```

EXploratory Graphs and Levene's test for equal variances

```{r}

hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~ common_name, scale = "free")

hists

qqs <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq(aes(color = common_name)) +
  facet_wrap(~ common_name, scale = "free")

qqs

# Based on qq-plots and histograms, observations for each group appear to be normally distributed. Also, n > 30 for all groups so according to CLT means will be normally distributed.

```

Let's evaluate 'dem variances and perform 'dat Levene's Test for equal variances homie!

```{r}

# If the largest sample varianes is < 4x greater than the smallest sample variance, then these are "cose enough" to consider variances equal

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

variances # The largest variance (33.1) is definitely not larger than 4x greater than the smallest variance (15.2)

# Levene's Test (since we have more than 2 variables)
#leveneTest function in the 'car' package

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)

levene_penguin

# H0: There is NO DIFFERENCE in variances; variances are EQUAL
# HA: Variances are NOT equal

# there is no significant difference in variances across species.


```

One-way ANOVA for penguin flipper lengths by species (common_name)

```{r}

# H0: The means accross all groups are equal
# HA: At least two group means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)

summary(penguin_aov)

# Post-hoc testing by Tukey's HSD


# H0: There is no significant difference in pairwise means
# HA: there IS a significant difference in pairwise means
penguin_ph <- TukeyHSD(penguin_aov)

penguin_ph

# aov.test function in the 'onewaytests' package DOES allow for inline references
```
Statement:
Mean flipper length of female penguins differed significantly for all species studied (one-way anova, F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05)
