---
title: "Penguin Flipper Lengths ANOVA"
author: "Geoffrey Cook"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load necessary packages and read in the necessary data

```{r}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
View(penguins)
```

Our question: Is there a difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipelago?

The type of test to compare means accross more than 2 groups for one indipendant variable...
ONE-WAY ANOVA

- Independent observations
- Equal variance (Levene's test)
- Normality (histograms/qqplots)

```{r}

counts <- with(penguins, table(species, sex))
View(counts)

# Let's do some updating!

penguins2 <- penguins %>% 
  mutate(
    common_name = 
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-species) %>% # REMOVES a particular column
  filter(sex == "FEMALE")
```

